---
title: "Homework 1"
author: "Rodrigo Petricioli"
date: "9/1/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
```


# Questions: 1.1, 1.2, 1.4, 1.5, 1.7

Note: Unless otherwise specified, all the numbers are referring to the corresponding exercises in our textbook: Statistical Computing with R (2nd Edition) by Maria L. Rizzo

## 1.1 
Generate a random sample $x_1 , ..., x_100$ of data from the $t4 (df=4)$ distribution using the $rt$ function. Use the $MASS::truehist$ function to display a probability histogram of the sample.

```{r echo=T}
set.seed(1234)
x <- rt(100,4)
MASS::truehist(x, col = "blue4", main = "t-Student Distribution")

```



## 1.2 
Add the $t4 density curve (dt)$ to your histogram in Exercise 1.1 using the curve function with $add=TRUE$.


```{r echo=T}
MASS::truehist(x, col = "blue4", main = "t-Student Distribution")
curve(dt(x, 4), add = T, col = "red4", lwd=4)


```

## 1.4 

a. Write an $R$ function $f$ in $R$ to implement the function 
  
  $f(x) = \frac{x-a}{b}$
    
that will transform an input vector $x$ and return the result. The function should take three input        arguments: $x, a, b$. 
  
b. To transform $x$ to the interval $[0,1]$ we substract the minimum value and divide by the range: 
  
  $y <- f(x, a = min(x), b = max(x) - min(x))$
  
Generate a random sample of $Normal(\mu = 2, \sigma = 2)$ data using $rnorm$ and use your function $f$ to transform this sample to the interval $[0,1]$. Print a summary of both the sample $x$ and the transformed sample $y$ to check the result. 



```{r echo=T}

f <- function(x, a, b){
  return (x-a)/b
}

x_norm <- rnorm(100,2,2)
y <- f(x_norm, min(x_norm), max(x_norm)-min(x_norm))

summary(x_norm)
summary(y)
```

## 1.5 

Refer to Exercise 1.4. Suppose that we want to transform the $x$ sample so that it has mean zero and standard deviation one (studentize the sample). That is, we want: 

  $z_i = \frac{x_i - \bar{x}}{s}, i = 1, ..., n$
  
where $s$ is the standard deviation of the sample. Using your function $f$ this is: 

  $z_i = f(x, a = mean(x), b = sd(x))$

Display a summary and histogram of the studentized sample $z$. It should be centered exactly ar zero. Use $sd(x)$ to check that the studentized sample has standard deviation exactly 1.0. 



```{r echo=T}
s <- f(x_norm, mean(x_norm), sd(x_norm))
z <- (x_norm - mean(x_norm))/sd(x_norm)
summary(z)
hist(z, col = "red4")
sd(z)
```

## 1.7
(ggplot) Refer to Example 1.14 where we displayed an array of scatterplots using $ggplot$ with $facet_wrap$. One of the variables in the $mpg$ data is $drv$, a character vector indicating whether the vehicle is front-wheel drive, rear-wheel drive, or four-wheel drive. Add $color = drv$ in $aes$:

  $aes(displ, hwy, color = drv)$
      
and display the revised plot. Your scatterplots should now have the three levels of $drv$ coded by color and the plot should have automatically generated a legend for $drv$ color.

```{r echo=T}
ggplot(mpg, aes(displ, hwy, color = drv)) + geom_point() + facet_wrap(~ class)

```
