---
title: Homework 3
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
set.seed(4444)
library('EnvStats')
library('VGAM')
mycol1 <- rgb(0, 0, 255, max = 255, alpha = 40, names = "blue50")
mycol2 <- rgb(200, 0, 0, max = 255, alpha = 40, names = "red50")
u <- runif(10000)
getmode <- function(x) {
  ux <- unique(x)
  ux[which.max(tabulate(match(x, ux)))]
}
```

## 3.3
The $Pareto(a,b)$ distribution has cdf $F(x) = 1 - (\frac{b}{x})^a,\ \  x\geq b>0,\ a>0.$ Derive the probability inverse transformation $F^{-1}(U)$ and use the inverse transformation method to simulate a random sample from the Pareto(2,2) distribution. Graph the density histogram of the sample with the Pareto(2,2) density superimposed for comparison. 

Ans. $$F^{-1}(U) = b(1-u)^{-1/a}$$

```{r echo=T}
set.seed(165)
n <- 1000
u <- runif(n) #Uniform Sample
a <- 2
b <- 2
X <- b*(1-u)^(-1/a) #Pareto Sample Using Inverse Transformation
p <- rpareto(n,2,2)
hist(X, freq=F, col=mycol1,breaks=200, xlim=c(0,15), main="Pareto(2,2)") #Histogram of Inverse Transformation Sample
hist(p, freq=F, col=mycol2,breaks=50, add=T, xlim=c(0,15)) #Histogram of Random Pareto Sample
curve(dpareto(x,2,2), add = T, col='green4', xlim=c(0,15)) #Density curve for Pareto(2,2)
```


## 3.4
The Rayleigh density is $f(x) = \frac{x}{\sigma^2} e^{-x^2/(2\sigma^2)},\ x\geq 0 ,\ \sigma>0.$ Develop an algorithm to generate random samples from a Rayleigh($\sigma$) samples for several choices of $\sigma > 0$ and check that the mode of the generated samples is close to the theoretical mode $\sigma$ (check the histogram). 

Ans.  
  
$F(x) = 1 - e^{-x^2 / 2\sigma^2}$ $=>$ $F^{-1}(U) = \sigma\sqrt{-2lnU}$ 

```{r echo=T}
raysample <- function(s, u){
  r <- s*sqrt(-2*log(u))
  return (r) 
  }

par(mfrow=c(1,3))

hist(raysample(1, u), freq=F, col='blue4', main="Rayleigh(1)")
curve(drayleigh(x,1), add=T, col='red4', lw=2) 
abline(v=1, col='green4')

hist(raysample(2, u), freq=F, col='blue4', main="Rayleigh(2)")
curve(drayleigh(x,2), add=T, col='red4', lw=2)
abline(v=2, col='green4')

hist(raysample(3, u), freq=F, col='blue4', main="Rayleigh(3)")
curve(drayleigh(x,3), add=T, col='red4', lw=2)
abline(v=3, col='green4')

```

```{r echo=F}
par(mfrow=c(1,1))
```

Table of the modes from the Rayleigh samples (rounded to one decimal) given different $\sigma$ values: 

|  $\sigma$  | Mode |  $\sigma$  | Mode |
|:---:|:---:|:---:|:---:|
|1|`r getmode(round(raysample(1, u),1))`| 5  |`r getmode(round(raysample(5, u),1))`| 
|2|`r getmode(round(raysample(2, u),1))`| 6  |`r getmode(round(raysample(6, u),1))`| 
|3|`r getmode(round(raysample(3, u),1))`| 7  |`r getmode(round(raysample(7, u),1))`| 
|4|`r getmode(round(raysample(4, u),1))`| 8  |`r getmode(round(raysample(8, u),1))`| 



## 3.5  
A discrete random variable $X$ has probability mass function    

|  x  |  0  |  1  |  2  |  3  |  4  |
|:---:|:---:|:---:|:---:|:---:|:---:|
|p(x)|0.1|0.2|0.2|0.2|0.3|

Use the inverse transform method to generate a random sample of size 1,000 from the distribution of $X$. Construct a relative frequency table and compare the empirical with the theoretical probabilities. Repeat using the $R\ sample$ function. 


```{r echo=T}

set.seed(165)

u2 <- runif(1000)

X2 <- numeric(1000)

for (i in 1:1000){
  
  if (u2[i] <= 0.1){
    X2[i] <- 0
  } else {
    if (u2[i] <= 0.3 && u2[i] >0.1){
      X2[i] <- 1
    } else{
      if( u2[i] <= 0.5 && u2[i] >0.3){
      X2[i] <- 2
      } else{
        if( u2[i] <= 0.7 && u2[i] >0.5){
        X2[i] <- 3
        } else{
          if( u2[i] <= 1 && u2[i] >0.7){
          X2[i] <- 4
          }
        }
      }
    }
  }
}

```


Comparative of Empirical Probabilities from the sample X and the Theoretical Probabilities: 
```{r echo=T}
df <- as.data.frame(round(table(X2)/1000,2))
df <- cbind(df, c(0.1,0.2,0.2,0.2,0.3))
colnames(df) <- c("x", "Empirical Prob.", "Theoretical Prob.")
df
```

Sample from X #1 (size = 100)
```{r echo=T}
set.seed(165)
X2_1 <- sample(X2, 100)

df1 <- as.data.frame(round(table(X2_1)/100,2))
df1 <- cbind(df1, c(0.1,0.2,0.2,0.2,0.3))
colnames(df1) <- c("x", "Empirical Prob.", "Theoretical Prob.")
df1
```

Sample from X #2 (size = 250)
```{r echo=T}
set.seed(165)
X2_2 <- sample(X2, 250)

df2 <- as.data.frame(round(table(X2_2)/250,2))
df2 <- cbind(df2, c(0.1,0.2,0.2,0.2,0.3))
colnames(df2) <- c("x", "Empirical Prob.", "Theoretical Prob.")
df2
```

Sample from X #3 (size = 500)
```{r echo=T}
set.seed(165)
X2_2 <- sample(X2, 500)

df2 <- as.data.frame(round(table(X2_2)/500,2))
df2 <- cbind(df2, c(0.1,0.2,0.2,0.2,0.3))
colnames(df2) <- c("x", "Empirical Prob.", "Theoretical Prob.")
df2
```


## 3.8
Write a function to generate random variates from a Lognormal($\mu, \sigma$) distribution using a transformation method, and generate a random sample  of size 1,000. Compare the histogram with the Lognormal density curve given by the $dlnorm$ function in $R$.

Ans. Given $X \sim N(\mu, \sigma)$ then $log(X) \sim LN(\mu, \sigma)$. Using this transformation from the $rnorm$ function in $R$. 

```{r echo=T}
set.seed(165)

normsample <- function(mu, sigma, n){
  return (rnorm(n,mu,sigma))
}


par(mfrow=c(1,3))

hist(exp(normsample(0,1,n)), col='blue4', freq=F, breaks=100, xlim=c(0,10), xlab = "LN(0,1) Sample", main="LogNormal(0,1)")
curve(dlnorm(x,0,1), col='red4', add=T)

hist(exp(normsample(2,1,n)), col='blue4', freq=F, breaks=100, xlim=c(0,50), xlab = "LN(2,1) Sample", main="LogNormal(2,1)")
curve(dlnorm(x,2,1), col='red4', add=T)
     
hist(exp(normsample(4,2,n)), col='blue4', freq=F, breaks=250, xlim=c(0,1500), xlab = "LN(4,2) Sample", main="LogNormal(4,2)")
curve(dlnorm(x,4,2), col='red4', add=T)

```

```{r echo=F}
par(mfrow=c(1,1))
```


## 3.9
The rescaled Epanechnikov kernel [92] is a symmetric density function $$f_e(x)=\frac{3}{4} (1-x^2),\ |x|\leq 1 $$ Devroye and Györfi give the following algorithm for simulation for this distribution. Generate $iid$ $U_1, U_2, U_3\ \sim Uniform(-1,1)$. If $|U_3|\geq|U_2|$ and $|U_3|\geq|U_1|$, deliver $U_2$; otherwise deliver $U_3$. Write a function to generate random variates from $f_e$, and construct the histogram density estimate of a large simulated random sample. 

```{r echo=T}
set.seed(555)
n2 <- 10000
U1 <- runif(n2,-1,1)
U2 <- runif(n2,-1,1)
U3 <- runif(n2,-1,1)
Uepa <- numeric(n2)

epa <- function(U1,U2,U3, n){
  Uepa <- numeric(n)  
  for (i in 1:n) {
    if (abs(U3[i]) >= abs(U2[i]) && abs(U3[i]) >= abs(U1[i])) {
      Uepa[i] <- U2[i]
      } else {
        Uepa[i] <- U3[i]
      }
  }
  return (Uepa)
}

hist(epa(U1,U2,U3, n2), col='blue4', freq=F, xlim=c(-1,1), main='Epanechnikov Distribution', xlab='Epanechnikov Estimate')
curve((3/4)*(1-x^2), add=T, col='red4', lw=2)
```





